---
title: "R coding Exercise"
output:
  word_document:
    toc: no
  html_document:
    toc: no
---

#First, load the "dslabs" package (install if needed). Also, load any other packages needed.
```{r}
library("dslabs")
library("tidyverse")
library(dplyr)
```

#Look at help file for gapminder data.
```{r}
help("gapminder") 
```

#Overview of data structure.
```{r}
str("gapminder")
```

#Summary of gapminder data.
```{r}
summary("gapminder")
```

#Determine object type of gapminder data.
```{r}
class("gapminder")
```

#Assigns only the African countries to a new object/variable called "africadata".
```{r}
africadata <- gapminder %>%
dplyr::filter(continent == "Africa")
```

#Check to see if you did it correclty (should have 2907 observations)
```{r}
str(africadata)
summary(africadata)
```
#Create new object with the Life expectancy and infant mortality of African Countries.
```{r}
africa_life_infant <- africadata %>%
  select("country", "life_expectancy", "infant_mortality")
```

#Create new object with the Life expectancy and population of African Countries.
```{r}
africa_life_pop <- africadata %>%
  select("country", "life_expectancy", "population")
```

#Plot life expectancy as a function of infant mortality.
```{r}
ggplot(data = africa_life_infant) + 
  geom_point(mapping = aes(infant_mortality, life_expectancy))
```

#Plot life expectancy as a function of infant mortality.
```{r}
ggplot(data = africa_life_pop) + 
  geom_point(mapping = aes(log10(population), life_expectancy))
```

#Identify missing Data for infant mortality.
```{r}
Missing1 <- africadata %>% 
           select(year,infant_mortality) %>% 
           filter(is.na(infant_mortality)) %>% 
           count(year) %>% 
           arrange(-n)
```

#Select data from Y2K
```{r}
Y2K_africadata <- africadata %>%
  filter(year == "2000")
```

#Create new object with the Life expectancy and infant mortality of African Countries for Y2K.
```{r}
Y2K_africa_life_infant <- Y2K_africadata %>%
  select("country", "life_expectancy", "infant_mortality")
```

#Create new object with the Life expectancy and population of African Countries.
```{r}
Y2K_africa_life_pop <- Y2K_africadata %>%
  select("country", "life_expectancy", "population")
```

#Plot life expectancy as a function of infant mortality.
```{r}
ggplot(data = Y2K_africa_life_infant) + 
  geom_point(mapping = aes(infant_mortality, life_expectancy))
```

#Plot life expectancy as a function of infant mortality.
```{r}
ggplot(data = Y2K_africa_life_pop) + 
  geom_point(mapping = aes(log10(population), life_expectancy))
```

#Linear Model: Infant Mortality, Life Expectancy for Y2K.
```{r}
fit_infant_life <- lm(life_expectancy~infant_mortality, data = Y2K_africa_life_infant)
summary(fit_infant_life)
```
#p-value = 2.826e-08; Based on the p-value, there is a significant correlation between infant mortality and life expectancy. As such, the slope of the relationship is negative, indicating a significant negative relationship between infant mortlaity and life expectancy. As infant mortality increases, life expectancy decreases.

#Linear Model: population, Life Expectancy for Y2K.
```{r}
fit_pop_life <- lm(life_expectancy~population, data = Y2K_africa_life_pop)
summary(fit_pop_life)
```
#P-value = 0.6159; Based on the p-value, no significant relationship between population size and life expectancy exists.